---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
llr_interpreter <- llr::llr_env$new()
knitr::knit_engines$set(
  clojure = llr::knitr_language_engine(llr_interpreter)
)
```

# llr

<!-- badges: start -->
<!-- badges: end -->

llr is a small, work in progress and just for fun clojure-like lisp on top of R's abstract syntax trees. Expressions are not interpreted, but are translated to R's AST and then interpreted by the R interpreter.

Most implementation details are sub-optimal, but the focus is on having fun and producing results instead writing perfect code. There are also many bugs and inconsistencies!

## Installation

``` r
remotes::install_github("dirkschumacher/llr")
```

## Intro

```{clojure}
(->
  r/datasets::mtcars
  (r/dplyr::filter (> hp 100))
  (r/dplyr::summarise :count (r/dplyr::n) :mean_mpg (r/mean mpg))
  (r/tibble::as_tibble))
```

Or run it from R

```{r}
library(llr)
interp <- llr_env$new()
interp$eval("(+ 1 1)")
```


## Special forms

### Data Types

#### Lists

```{clojure}
; this is a list
'(1 2 3 4 5 6)
; an unquoted list is a function call
(+ 1 2 3 4 5 6)
```

#### Vectors

```{clojure}
[1 2 3 4]
```

#### Maps

```{clojure}
{:a 1 :b 2}
```


### Symbols

```{clojure, eval = FALSE}
x
```

```{clojure, eval = FALSE}
namespaced.variable/x
```

```{clojure, eval = FALSE}
:keyword
```

```{clojure, eval = FALSE}
"character"
```

```{clojure, eval = FALSE}
10 ; integer
```

```{clojure, eval = FALSE}
10.42 ; double
```

### Functions

```{clojure, eval = FALSE}
(fn [a b] (+ a b))

(fn this
  ([] 0)
  ([a] a)
  ([a b] (+ a b))
  ([a b & more] (reduce + (conj [a b] more))))
```

### def

To define symbols something within a namespace use `def`

```{clojure}
(def x 1)
(def plus (fn [a b] (+ a b)))
(plus x x)
```


### Meta-data

Symbols and values can hold meta-data. That meta-data needs to be a map at the moment.

```{clojure}
(def ^{:const true} x ^{:meta "hello"} [ 1 2 3])
(meta x)
```

Meta-data on symbols is currently only available to the reader.

### Macros

Macros are also supported. Macros are functions bound to a name with meta data `{:macro true}`.

In a macro you can use syntax-quote `<backtick>` together with the unquote `~` and unquote-splice `~@` operators.

```{clojure}
(defmacro infix [operand1 operator operand2]
  `(~operator ~operand1 ~operand2))
(infix 1 + 1)
```


### Recursion

Similar to Clojure llr uses `recur` to jump to a recursion point currently only defined by `loop`.

```{clojure}
(def is-even 
  (fn [number] 
    (loop [cnt number]
      (if (zero? cnt)
        true
        (if (< cnt 0) false (recur (- cnt 2)))))))
```

```{clojure}
(is-even 5001)
```

```{clojure}
(is-even 5000)
```

### Namespaces

Every top level definition is part of a namespace

```{clojure}
(ns product.lib)
(defn compute [a b] (+ a b))
(ns user)
(product.lib/compute 10 32)
```


### Reader Dispatch

The reader switches to a different set of interpretations of the next symbol when reading the character `#`.

### `#_` ignores the next form

```{clojure}
#_ (r/stop "error")
"Yay"
```

### R interop

All symbols starting with the namespace `r/` are treated slightly different. You can use that to refer to external R functions and symbols. In addition keywords are interpreted as named arguments.

```{clojure}
(r/set.seed 1)
(def rand-numbers (r/stats::rnorm :n 10))
(r/mean rand-numbers)
```

## Design Goals

* Have fun, experiment and learn :)
* Build a clojure-like language that supports R-interop using the `r/` namespace.

## Contributing

* Please read the code-of-conduct and also be aware that this a fun project,
  so things will break and progress is valued prefect code (at the moment).
* However everyone is invited to play around with the language, learn together,
  extend it, fix bugs and propose features.

## Code of Conduct

Please note that the llr project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
